---
title: "School rankings with TGVE"
output: github_document
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(dirname(inputFile),'/README.md')) })
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

Let us show how the TGVE can be used to see rankings of schools in the UK with fewest steps possible. For the purpose of this Rmarkdown doc, I will create a directory at:

```{r}
library(readr)
library(dplyr)
d = "~/code/tgve/eatlas-data/schools/data"
if(!dir.exists(d)) {
  dir.create(d)
}
```


## Get school data
The school data is availabe from https://get-information-schools.service.gov.uk/Downloads. For the purpose of this analysis download "All establishment data" by checking "Establishment fields CSV, 60.56 MB" checkbox. If you download that you will get a file with name `dubasealldata20220306.csv`.

You if you then read the contents of that CSV with `readr` you would be looking at something like this:

```{r}
csv = readr::read_csv("~/Downloads/edubasealldata20220306.csv")
head(names(csv))
names(csv)[grep("ofsted", names(csv), ignore.case = TRUE)]
# lots of code columns
head(names(csv)[grep("code", names(csv), ignore.case = TRUE)])
# this is what we will be using
names(csv)[grep("district", names(csv), ignore.case = TRUE)]

```
So let us get only the ranking and the geography as:

```{r}
data = csv %>% select(`OfstedRating (name)`, `DistrictAdministrative (code)`, `DistrictAdministrative (name)`)
names(data) = c("ranking", "la_code", "name")
# we can remove the na's in the csv like
# data = data %>% filter(!is.na(ranking))
data %>% group_by(ranking, la_code, name) %>% summarise(ranking_count = n()) %>% 
  readr::write_csv(file.path(d, "data.csv"))
```

## Get geography
The reason I named second column `la_code` is that I know we have that Local Authority District geographies over at SaferActive project here. We can of course download the file and save it in the same directory:

```{r}
geoURL = "https://raw.githubusercontent.com/saferactive/tgve/main/las-only-code.geojson"
download.file(geoURL, file.path(d, "lad.geojson"))
```

Now, feel free to explore this using `sf` or any other way you do your geoprocessing but there will be a column named `la_code` as `tgver` will need a matching column name to put the two files together.

## `tgver::explore_dir`
Make sure you have `tgver`

```{r, eval=FALSE}
install.packages("tgver")
```

Step three:
```{r, eval=FALSE}
tgver::explore_dir(d)
```

We can do even better, all we have to do is get the data somewhere, so I added it to the `tgve/example-data` repository, therefore we can do:

```{r final, out.width="100%"}
# the link to browse
# https://tgve.github.io/app/?defaultURL=https://raw.githubusercontent.com/tgve/example-data/main/schools/data.csv&geographyURL=https://raw.githubusercontent.com/saferactive/tgve/main/las-only-code.geojson&geographyColumn=la_code
dataURL = "https://raw.githubusercontent.com/tgve/example-data/main/schools/data.csv"
app = "https://tgve.github.io/app/"
knitr::include_url(paste0(
  app, "?defaultURL=", dataURL, "&geographyURL=", geoURL, "&geographyColumn=la_code"
))
```


